---
output:
  html_document: default
  pdf_document: default
---
# Geographic Data in R (Chapter 2: Geocomputations in R)

<https://geocompr.robinlovelace.net/spatial-class.html#crs-intro>

## Install packages

```{r}
install.packages("sf")
install.packages("raster")
install.packages("spData")
remotes::install_github("Nowosad/spDataLarge")
```
## Load libraries

```{r}
library(sf)
library(raster)
library(spData)
library(spDataLarge)
# Load some vignettes
vignette(package = "sf")
vignette("sf1")
```

# Vector Data
## Load the `world` data set from spData

```{r}
names(world) #last column contains geographic info
world$geom
summary(world["lifeExp"])
world_mini = world[1:2, 1:3]
world_mini
# Plotting multiple variables
plot(world[3:6])
# Plotting a single variable
plot(world["pop"])
```
## Add maps as layers

```{r}
world_asia = world[world$continent == "Asia", ]
asia = st_union(world_asia)
plot(world["pop"], reset = FALSE)
plot(asia, add = TRUE, col = "red")
```

## Converting a polygon to point with st_centroid

```{r}
plot(world["continent"], reset = FALSE)
cex = sqrt(world$pop) / 10000
world_cents = st_centroid(world, of_largest = TRUE)
plot(st_geometry(world_cents), add = TRUE, cex = cex)
```

## Add geographic features like a window expansion

```{r}
india = world[world$name_long == "India", ]
plot(st_geometry(india), expandBB = c(0, 0.2, 0.1, 1), col = "gray", lwd = 3)
plot(world_asia[0], add = TRUE)
```

Simple feature objects:

* A point: st_point()
* A linestring: st_linestring()
* A polygon: st_polygon()
* A multipoint: st_multipoint()
* A multilinestring: st_multilinestring()
* A multipolygon: st_multipolygon()
* A geometry collection: st_geometrycollection()


## sf object structure

### we will represent a temperature of 25°C in London on June 21st, 2017

```{r}
lnd_point = st_point(c(0.1, 51.5))                 # sfg object (simple feature geometry)
lnd_geom = st_sfc(lnd_point, crs = 4326)           # sfc object (simple feature column)
lnd_attrib = data.frame(                           # data.frame object
  name = "London",
  temperature = 25,
  date = as.Date("2017-06-21")
  )
lnd_sf = st_sf(lnd_attrib, geometry = lnd_geom)    # sf object
lnd_sf
class(lnd_sf)
```
# Raster Data

### Read data from a file

```{r}
help("raster-package")
# Load Zion National park digital elevation model
raster_filepath = system.file("raster/srtm.tif", package = "spDataLarge")
new_raster = raster(raster_filepath)
new_raster
plot(new_raster)
```


## Creat a raster file from scratch

```{r}
new_raster2 = raster(nrows = 6, ncols = 6, res = 0.5,xmn = -1.5, xmx = 1.5, ymn = -1.5, ymx = 1.5, vals = 1:36)
```


## Multiple raster layers

```{r}
multi_raster_file = system.file("raster/landsat.tif", package = "spDataLarge")
r_brick = brick(multi_raster_file)
r_brick
nlayers(r_brick)
```


# Coordinate Reference Systems

* Geographic coordinate system: Latitude and longitude. Earth represented by an spherical or ellipsoidal surface.
* Two main ways to describe CRS in R are an `epsg` code or a `proj4string` definition
* Projected CRSs are based on Cartesian coordinates on an implicitly flat surface. They have an origin, x and y axes, and a linear unit of measurement such as meters. 

```{r}
crs_data = rgdal::make_EPSG()
View(crs_data)
vector_filepath = system.file("vector/zion.gpkg", package = "spDataLarge")
new_vector = st_read(vector_filepath)
st_crs(new_vector) # get CRS
projection(new_raster) # get CRS from a raster
```

